### 2016-01-15 04-evaluation.tex
A table w/ breakdown for the pipelining

In order to evaluate the horizontal scalability of the distributed
implementation we tested the system's performance across different cluster
sizes while holding the problem size constant. For this study, we used the
com-Friendster graph as it contains the largest number of vertices and edges.
%
Figure~\ref{fig-strong-scaling}-a presents the execution time of 2048 algorithm
iterations across multiple cluster sizes. The x-axis starts from 8 worker nodes
as the data set is too large to fit into the collective memory of a smaller
cluster.
%
As shown in the figure, the execution
time steadily decreases by increasing the cluster size.
A deeper analysis of the individual execution phases of the algorithm provides
insights into the scalability if its building blocks.
%
In addition to the total execution time, Figure~\ref{fig-strong-scaling}-a
presents the cumulative time spent in individual computational phases across
iterations. Moreover, Figure~\ref{fig-strong-scaling}-b presents the speedup
achieved for the same experiments reported in
Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
the execution is update\_pi.
The reported total time for each cluster size is significantly less than the
sum of the execution times of the individual phases. This is due to the
overlapping execution of the two most expensive phases, namely, update\_pi and
deploy\_minibatch. Both of these phases initially gain significant speedup with the
addition of compute nodes. However, the speedup curve gradually slows down for
larger cluster sizes as the work granularity of each worker node decreases,
limiting their resource utilization.
%
The time spent in update\_beta remains relatively constant across cluster sizes
as it performs an insignificant amount of work compared to the synchronization
overhead of a collective MPI reduction operation contained within it.

@@

>>>>

In order to evaluate the horizontal scalability of the distributed
implementation we tested the system's performance across different cluster
sizes while holding the problem size constant. For this study, we used the
com-Friendster graph as it contains the largest number of vertices and edges.
%
Figure~\ref{fig-strong-scaling}-a presents the execution time of 2048 algorithm
iterations across multiple cluster sizes. The x-axis starts from 8 worker nodes
as the data set is too large to fit into the collective memory of a smaller
cluster.
%
As shown in the figure, the execution
time steadily decreases by increasing the cluster size.
A deeper analysis of the individual execution phases of the algorithm provides
insights into the scalability if its building blocks.
%
In addition to the total execution time, Figure~\ref{fig-strong-scaling}-a
presents the cumulative time spent in individual computational phases across
iterations. Moreover, Figure~\ref{fig-strong-scaling}-b presents the speedup
achieved for the same experiments reported in
Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
the execution is update\_pi.
The reported total time for each cluster size is significantly less than the
sum of the execution times of the individual phases. This is due to the
overlapping execution of the two most expensive phases, namely, update\_pi and
deploy\_mini-batch. Both of these phases initially gain significant speedup with the
addition of compute nodes. However, the speedup curve gradually slows down for
larger cluster sizes as the work granularity of each worker node decreases,
limiting their resource utilization.
%
The time spent in update\_beta remains relatively constant across cluster sizes
as it performs an insignificant amount of work compared to the synchronization
overhead of a collective MPI reduction operation contained within it.

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{l d{5.1} d{5.1}}
    \textbf{Iteration stages}  & \multicolumn{1}{c}{non-pipelined}
                                          & \multicolumn{1}{c}{pipelined} \\
    \hline
    total                      & 450    & 365 \\
    draw/deploy mini-batch      &  45.6  &      \\
    update\_phi                & 285    & 241 \\
    update\_pi                 &   3.8  &   4.6 \\
    update $\beta/\theta$      &  25.9  &   33.6 \\
    \hline
    \textbf{Substages within update\_phi} \\
    draw/deploy mini-batch      &        &  26.2 \\
    load $\pi$                 & 205    & 209 \\
    update $\phi$              &  74    &  74 \\
    \hline
  \end{tabular}
  \caption{The most important stages in the execution; com-Friendster on 65
  compute nodes, with 12K communities. Times in ms per iteration. The lower
  half of the table shows substages within \textit{update\_phi}. In the
  pipelining version, these components are done in parallel.}
  \label{table-pipeline}
\end{table}

Table~\ref{table-pipeline} shows a breakdown of the most important time
consumers; the left-hand column is without pipelining, the right-hand
column has pipelining. \textit{update\_phi} is the dominant contribution,
and within that, loading $\pi$ from the DKV store takes by far the most time.
The pipelining optimization overlaps loading $\pi$ with generation and
deployment of the mini-batch, and with the calculation of \textit{update\_phi}.
